### 2016-06-13 capability-machine.tex
small changes and typos

\newcommand{\RegionName}{\plaindom{RegionName}}
\newcommand{\Worlds}{\plaindom{World}}
\newcommand{\UPred}[1]{\plaindom{UPred}(#1)}

@@

 Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.

@@

\section{Logical Relation}
\label{sec:logical-relation}
\subsection{Worlds}
\[
\Worlds \cong \frac{1}{2} (\RegionName \parfun (\Worlds \monnefun \UPred{\Addrs \parfun \Words}))
\]

\[
  \iota_{\var{start},\var{end}} \; : \; \Worlds \monfun \UPred{\Addrs \parfun \Heaps}
\]
\begin{align*}
  \iota_{\var{start},\var{end}}(W) \defeq & \{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                     & \quad \forall \addr \in \dom(\heapseg) \ldotp \npair{\heapseg(\addr)} \in \stdvr(W) \}
\end{align*}

\begin{align*}
  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionName \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionName \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
  \execCond{n,W,\start,\addrend,\perm} =  & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
                                          & \; \qquad \npair{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
\end{align*}

\[
\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
\]
\begin{align*}
  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
\union \\
                   & \{ \npair{\stdcap[\noperm] }  \} 
\union \\
                   & \{ \npair{\stdcap[\readonly] } \mid \readCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\readwrite] } \mid \writeCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\exec]} \mid \\
                   & \quad\readCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\exec} \}
\union \\
% For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
                   & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\rwx] = c} \mid \\
                   & \quad \writeCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\exec} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\rwx} \}
\end{align*}

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\
  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid \\
                           & \quad (\exists \reg', \heap' \ldotp (\reg,\heap) \step[n] (\reg',\heap')) \lor \\ %Define the step stuff 
                           & \quad (\exists \reg', \heap', i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
                           & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap)  \\
                           & \qquad \Rightarrow \npair[n-i]{\heap(0)} \in \stdvr(W) \land \\
                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap \restrictfun{\{0\}}} \in W(r)(W) )\}
\end{align*}
The ``good observations'' are as follows
\begin{itemize}
\item The execution does not stop in $n$ steps. 
\item The execution stops with a $\failed$.
\item The execution halts in at most $n$ steps with a valid value in in address 0 and the value at address 0 is also valid with respect to the world.
\end{itemize}
Address 0 is the designated I/O address.

>>>>

\newcommand{\RegionName}{\plaindom{RegionName}}
\newcommand{\Worlds}{\plaindom{World}}
\newcommand{\StorePred}{\plaindom{StorePred}}
\newcommand{\UPred}[1]{\plaindom{UPred}(#1)}

@@

 Given $\Phi$, if $\memheap$ contains the ticket dispenser program starting at the address labelled $\init$, $\memreg(\pcreg) = (\rwx,\init,\init+\var{proglen}-1,\init)$, $\memreg(\rio) = (\readwrite,0,0,0)$, the remaining registers of $\memreg$ contains 0, and $\Phi \rightarrow^* (\halted,\Phi')$, then the I/O address (that is $\memheap[\Phi'](0)$) contains either $-1$ or a number that is $\geq 0$ and even.

@@

\section{Logical Relation}
\label{sec:logical-relation}
\subsection{Worlds}
\[
  \StorePred = \Worlds \monnefun \UPred{\Addrs \parfun \Words}
\]
\[
\Worlds \cong \frac{1}{2} (\RegionName \parfun \StorePred)
\]

\[
  \iota_{\var{start},\var{end}} \; : \; \StorePred
\]
\begin{align*}
  \iota_{\var{start},\var{end}}(W) \defeq & \{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                     & \quad \forall \addr \in \dom(\heapseg) \ldotp \npair{\heapseg(\addr)} \in \stdvr(W) \}
\end{align*}

\begin{align*}
  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionName \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionName \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
  \execCond{n,W,\start,\addrend,\perm} =  & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
                                          & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \\
\end{align*}

\[
\stdvr \; : \;  \Worlds \monfun \UPred{\Words}
\]
\begin{align*}
  \stdvr(W) \defeq & \{ \npair{i} \mid i \in \ints \} 
\union \\
                   & \{ \npair{\stdcap[\noperm] }  \} 
\union \\
                   & \{ \npair{\stdcap[\readonly] } \mid \readCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\readwrite] } \mid \writeCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\exec]} \mid \\
                   & \quad\readCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\exec} \}
\union \\
% For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
                   & \{ \npair{\stdcap[\entry]} \mid \entryCond{n,W,\start,\addrend} \} \union \\
                   & \{ \npair{\stdcap[\rwx]} \mid \\
                   & \quad \writeCond{n,W,\start,\addrend} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\exec} \land \\
                   & \quad \execCond{n,W,\start,\addrend,\rwx} \}
\end{align*}

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Confs} \\
  \observations (W) \defeq & \{ \npair{(\reg,\heap)} \mid \\
                           & \quad (\exists \reg', \heap' \ldotp (\reg,\heap) \step[n] (\reg',\heap')) \lor \\ %Define the step stuff 
                           & \quad (\exists \reg', \heap', i \leq n \ldotp (\reg,\heap) \step[i] \failed) \lor \\
                           & \quad (\forall \heap', i \leq n \ldotp (\reg,\heap) \step[i] (\halted,\heap')  \\
                           & \qquad \Rightarrow \npair[n-i]{\heap'(0)} \in \stdvr(W) \land \\
                           & \qquad \qquad \forall r \ldotp 0 \in \dom(W(r)) \ldotp \npair[n-i]{\heap' \restrictfun{\{0\}}} \in W(r)(W) )\}
\end{align*}
The ``good observations'' are as follows
\begin{itemize}
\item The execution does not stop in $n$ steps. 
\item The execution stops with a $\failed$.
\item The execution halts in at most $n$ steps with a valid value in in address 0 and the value at address 0 is also valid with respect to the world.
\end{itemize}
Address 0 is the designated I/O address.
